---
title: "Survival"
output:
  pdf_document:
    toc: yes
    toc_depth: '4'
  html_document:
    toc: yes
    toc_depth: 4
    theme: united
    highlight: tango
---


# Data Dictionary 

Survival in a randomised trial comparing two treatments for ovarian cancer


futime:	  survival or censoring time

fustat:	  censoring status

age:	    in years

resid.ds:	residual disease present (1=no,2=yes)

rx:	      treatment group

ecog.ps:	ECOG performance status 

ECOG: 

https://ecog-acrin.org/resources/ecog-performance-status 

  1 - Restricted in physically strenuous activity but ambulatory and 
  
  able to carry out work of  a light or sedentary nature, e.g., 
  
  light house work, office work     
  
  2 - Ambulatory and capable of all selfcare but unable to carry out 
  
  any work activities; up and about more than 50% of waking hours

# Packages

For package installation we used function from - <https://stackoverflow.com/a/31620488>. 

It will check for missing packages and install them + it will query both - CRAN and Bioconductor repositories for missing package.

```{r}
install.packages.auto <- function(x) { 
  x <- as.character(substitute(x)) 
  if(isTRUE(x %in% .packages(all.available=TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else { 
    #update.packages(ask= FALSE) #update installed packages.
    eval(parse(text = sprintf("install.packages(\"%s\", dependencies = TRUE)", x)))
  }
  if(isTRUE(x %in% .packages(all.available=TRUE))) { 
    eval(parse(text = sprintf("require(\"%s\")", x)))
  } else {
    source("http://bioconductor.org/biocLite.R")
    #biocLite(character(), ask=FALSE) #update installed packages.
    eval(parse(text = sprintf("biocLite(\"%s\")", x)))
    eval(parse(text = sprintf("require(\"%s\")", x)))
  }
}

```


## Install 

```{r}
install.packages.auto("survival")
install.packages.auto("dplyr")
install.packages.auto("ggplot2")
install.packages.auto("tidyr")
install.packages.auto("ggpubr")
install.packages.auto("vcd")
install.packages.auto("corrplot")
install.packages.auto("survminer")
install.packages.auto("ggfortify")
```

install.packages("expss")


## Library

```{r}
library(survival)
library(dplyr)
library(ggplot2)
library(tidyr)

library(ggpubr)
library(vcd)
library(corrplot)
library(survminer)
library(ggfortify) 

```
## Plot themes

```{r}
theme_set(theme_pubr())
```


# Data analise

## EDA

```{r}
df_raw <- (ovarian)
df <- (ovarian)
head(df)
```


### NA 

```{r}
colSums(is.na(df))
```

No values in daatset.

### Check data type

```{r}
str(df)
summary(df)
```

```{r}
# df$fustat <- as.factor(df$fustat)
df$resid.ds <- factor(df$resid.ds, levels = c("1", "2"), labels = c("No", "Yes"))
df$rx <- factor(df$rx,  levels = c("1", "2"), labels = c("A", "B"))
df$ecog.ps <- factor(df$ecog.ps, levels = c("1", "2"), labels = c("I", "II"))

str(df)
```

### General statistics 

```{r}
summary(df)
```

Dataset:

  * There are data on 26 patients
  
  * Out of 26 patients, for 14 data are censored
  
  * Residual disease is not present in 15 patients and present in 11 patients
  
  * 14 patients can do only light sedentary work and restricted in physically strenuous 
  
  * and 12 patients capable of all selfcare but unable to carry out any work activities
  
  * Patients groups divided by type of Treatment type are balanced (A: 13 and B: 13)
  
  * The youngest patient - 38, the oldest patient - 74
  
  * Minimum patient follow-up time - 59, maximum patient follow-up time - 1227


### Mosaicplot

This graph clearly demonstrates the number of patients in groups

distinguished according to the characteristics: type of treatment (rx), 

presence/absence of residual disease ("resid.ds), ECOG status (ecog.ps), 

is the data censored at the moment (fustat).


```{r}
mosaic( ~ rx + fustat + ecog.ps + resid.ds, data = df,
         highlighting = "resid.ds", 
        highlighting_fill = c("pink", "lightblue"),
        labeling = labeling_values,
        main = "Dataset subgroups",
        sub = "Red - residual disease present, Blue residual disease not present")
```

### Barplots and boxplots

```{r}

age_boxplot <- ggboxplot(df, x = "rx", y = "age", 
          ylab = "Age", 
          xlab = "Treatment group",
          color = "rx", 
          palette = "jco",
          add = "dotplot",  
          add.params = list(dotsize = 0.6),
          rotate = TRUE)


age_barplot <- ggplot(df, aes(x = rx, y = age, fill = age)) + 
  geom_bar(stat = "identity") +
  coord_flip() + 
  scale_fill_gradient(low='yellow', high='slateblue4', space='Lab') + 
  xlab("") + 
  ylab("Age") + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="right")

futime_boxplot <- ggboxplot(df, x = "rx", y = "futime", 
          ylab = "Futime", 
          xlab = "Treatment group",
          color = "rx", 
          palette = "jco",
          add = "dotplot",  
          add.params = list(dotsize = 0.6),
          rotate = TRUE)


futime_barplot <- ggplot(df, aes(x = rx, y = futime, fill = futime)) + 
  geom_bar(stat = "identity") +
  coord_flip() + 
  scale_fill_gradient(low='yellow', high='slateblue4', space='Lab') + 
  xlab("") + 
  ylab("Futime") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        legend.position="right")



fig_1 <- ggarrange(age_boxplot, age_barplot, 
                   futime_boxplot, futime_barplot,
                   labels = c("A", "B", "C", "D"),
                   ncol = 2, nrow = 2)
fig_1


```


* The average age of patients treated with different drugs is approximately the same.However, the range of ages in group A is much greater than in group B (Pic. A-B)

* The average follow-up time for patients taking different treatments varies slightly. Patients from group B, in which the range of ages is less, were observed on average a little longer


### Correlation

Based on http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram


To compute the matrix of p-value:

```{r}
# mat : is a matrix of data
# ... : further arguments to pass to the native R cor.test function
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}
# matrix of the p-value of the correlation
p.mat <- cor.mtest(df_raw)
head(p.mat)
```

```{r}
col <- colorRampPalette(c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"))


corrplot(cor(df_raw), method="color", col=col(200),  
         type="lower", order="hclust", 
         addCoef.col = "black", # Add coefficient of correlation
         tl.col="black", tl.srt=45, #Text label color and rotation
         # Combine with significance
         p.mat = p.mat, sig.level = 0.05, insig = "blank", 
         # hide correlation coefficient on the principal diagonal
         diag=FALSE 
         )
```


* Positively correlated with each other - age of the patient and data censored or not at the moment. 

* Negatively correlated with each other - data censored or not at the moment and patient follow-up time. 

* Also negatively correlated with each other - patient follow-up time and age. 

For some reason, older patients were seen less by researchers.


## Kaplan–Meier + logrank_test

```{r}
km <- with(df, Surv(futime, fustat))
km
```

### Survival factor based on time

```{r}
km_fit <- survfit(Surv(futime, fustat) ~ 1, data = df)

autoplot(km_fit) +
  labs(title = "Survival factor based on time")

```

### Survival factor based on time + Treatment type

```{r}
km_rx_fit <- survfit(Surv(futime, fustat) ~ rx, data=df)
summary(km_rx_fit)


autoplot(km_rx_fit) 


survdiff(Surv(df$futime, df$fustat) ~ rx, data = df)
logrank_test(Surv(df$futime, df$fustat) ~ rx, data = df)


```

* Patients treated according to protocol "B" had a longer follow-up time (survival rate is higher)


### Survival factor based on time + Age

```{r}
df_new <- df %>% mutate(age_factor = ifelse(age >=56, ">56", "<56"))
df_new$age_factor <- as.factor(df_new$age_factor)


```

```{r}
km_age <- survfit(Surv(futime, fustat) ~ age_factor, data=df_new)

summary(km_age)

autoplot(km_age) 


survdiff(Surv(df$futime, df$fustat) ~ age_factor, data = df_new)
logrank_test(Surv(df$futime, df$fustat) ~ age_factor, data = df_new)

```

* Patients whose age is below average (<56) had a longer follow-up (survival rate is higher)

### Survival factor based on time + ECOG status


```{r}
km_ecog <- survfit(Surv(futime, fustat) ~ ecog.ps, data=df)

autoplot(km_ecog) 

summary(km_ecog)


survdiff(Surv(df$futime, df$fustat) ~ ecog.ps, data = df)
logrank_test(Surv(df$futime, df$fustat) ~ ecog.ps, data = df)

```


* Patients in worse condition (ECOG II) had a shoter follow-up time (survival rate is lower)


### Survival factor based on time + resid.ds


```{r}
km_resid_ds <- survfit(Surv(futime, fustat) ~ resid.ds, data=df)

autoplot(km_resid_ds) 

summary(km_resid_ds)


survdiff(Surv(df$futime, df$fustat) ~ resid.ds, data = df)
logrank_test(Surv(df$futime, df$fustat) ~ resid.ds, data = df)

```

* Patients with residual disease had a shoter follow-up time (survival rate is lower)



### Cox

```{r}
cox <- coxph(formula = Surv(futime, fustat) ~ age_factor  + rx + resid.ds + ecog.ps, data = df_new)
summary(cox)

```
```{r}
cox_fit <- survfit(cox)
plot(cox_fit)
```


```{r, message=FALSE}
aa_fit <- aareg(formula = Surv(futime, fustat) ~ age_factor + rx + resid.ds + ecog.ps, data = df_new)
autoplot(aa_fit)
```


## Hazard Ratio

```{r}
fit.coxph <- coxph(Surv(futime, fustat) ~ age_factor + rx + resid.ds + ecog.ps, data = df_new)
ggforest(fit.coxph, data = df_new)

```

** Younger patients, without residual disease, in better condition (ECOG II) and treated according to protocol "B", have better chances of survival.*

